---
title: "R Visualization with ggplot"
description: |
  Bascic ggplot using R
author:
  - name: Linya Huang
    url: https://www.linkedin.com/in/linya-huang/
date: 07-03-2021
output:
  distill::distill_article:
    toc: true
    toc_depth: 6
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message = FALSE)
```

## Install and Launch R packages

Insert an R code chunk either using the RStudio toolbar (the Insert button) or the keyboard shortcut Ctrl + Alt + I (Cmd + Option + I on macOS).


```{r, include=TRUE,message = FALSE}
packages= c('DT','ggiraph','plotly','tidyverse','DT')

for(p in packages){
  if(!require(p,character.only= T)){
    install.packages(p)
    }
  library(p, character.only = T)
}
```

## Importing Data

```{r}
exam_data = read_csv("data/Exam_data.csv")
summary(exam_data) #summary statistics of data 
```

## [ggplot](https://ggplot2.tidyverse.org/reference/ggplot.html)

Handy Manual

[Cheat Sheet](https://github.com/rstudio/cheatsheets/blob/master/data-visualization-2.1.pdf)


### Bar Chat of discrete values __geom_bar__

```{r}
ggplot(exam_data, 
       aes(x=RACE))+
  geom_bar()
```

### Distribution of maths score using __geom_dotplt__

```{r}
ggplot(exam_data, 
       aes(x=MATHS))+
  geom_dotplot(binwidth=2.5,
               dotsize = 0.5)+
  scale_y_continuous(NULL,
                     breaks=NULL)
```

```{r}
ggplot(exam_data, 
       aes(x=MATHS,
           fill=RACE))+
  geom_dotplot(binwidth=2.5,
               dotsize = 0.5)+
  scale_y_continuous(NULL,
                     breaks=NULL)
```


### Scatter Plot __geom_jitter__

```{r, error=FALSE}
ggplot(exam_data, 
       aes(x=MATHS,
           y=ENGLISH))+
  geom_jitter(height=2,
              width = 2)+
  xlim(0, 100)+ylim(0, 100)+
  geom_abline(coef = c(0,1))
```

### Histogram __geom_histogram__

```{r}
ggplot(exam_data, 
       aes(x=MATHS))+
  geom_histogram(bins=20,
                 color="black",
                 fill="light blue")
```

Fill with gender and change the interior color 

```{r}
ggplot(exam_data, 
       aes(x=MATHS,
           fill=GENDER))+
  geom_histogram(bins=20,
                 color="grey30")
```

### Box Plot __geom_boxplot__

```{r}
ggplot(exam_data, 
       aes(y=MATHS,
           x=GENDER,color=GENDER))+
  geom_boxplot()+
  geom_point(position = "jitter",
             size=1)
```



### Interactive __[ggiraph](https://davidgohel.github.io/ggiraph/)__


__ggiraph__ makes the chart more interactive, hover to the dot and show toolip

```{r}
p = ggplot(exam_data, 
       aes(x=MATHS))+
  geom_dotplot_interactive(
    aes(tooltip = ID),
    stackgroups = TRUE,
    binwidth=1,
    method="histodot")+
    scale_y_continuous(NULL,
                       breaks = NULL)
girafe(
  ggobj = p,
  width_svg = 6,
  height_svg = 6*0.618
  )
```

Set aes(data_id=CLASS,tooltip=CLASS), to check other classmates

```{r}
p = ggplot(exam_data, 
       aes(x=MATHS))+
  geom_dotplot_interactive(
    aes(data_id=CLASS,
        tooltip=CLASS),
    stackgroups = TRUE,
    binwidth=1,
    method="histodot")+
    scale_y_continuous(NULL,
                       breaks = NULL)
girafe(
  ggobj = p,
  width_svg = 6,
  height_svg = 6*0.618
  )
```
# [plotly](https://plotly.com/r/)

## Basic

```{r}

pal <- c("purple","blue","grey","yellow")
plot_ly(data=exam_data,
        x=~MATHS,
        y=~ENGLISH,
        color = ~RACE,
        colors = pal)
```

Tooltips
```{r}

pal <- c("purple","blue","grey","yellow")
plot_ly(data=exam_data,
        x=~MATHS,
        y=~ENGLISH,
        text= ~paste("Student ID:",ID,
                     "<br>Class:",CLASS),
        color = ~RACE,
        colors = pal)
```

Layout

```{r}

pal <- c("purple","blue","grey","yellow")
plot_ly(data=exam_data,
        x=~MATHS,
        y=~ENGLISH,
        text= ~paste("Student ID:",ID,
                     "<br>Class:",CLASS),
        color = ~RACE,
        colors = pal) %>%
  layout(title="English Score versus Math Score",
         xaxis=list(range=c(0,100)),
         yaxis=list(range=c(0,100)))
```



ggplotly

```{r, error=FALSE}
p<-ggplot(exam_data, 
       aes(x=MATHS,
           y=ENGLISH))+
  geom_point(dotsize=1)+
  xlim(0, 100)+ylim(0, 100)+
  geom_abline(coef = c(0,1))

ggplotly(p)
```
## Coordinated Multiple Views with **[plotly](https://plotly.com/r/subplots/)**


```{r}
p1<-ggplot(exam_data, 
       aes(x=MATHS,
           y=ENGLISH))+
  geom_point(dotsize=1)+
  xlim(0, 100)+ylim(0, 100)+
  geom_abline(coef = c(0,1))

p2<-ggplot(exam_data, 
       aes(x=MATHS,
           y=SCIENCE))+
  geom_point(dotsize=1)+
  xlim(0, 100)+ylim(0, 100)+
  geom_abline(coef = c(0,1))

subplot(ggplotly(p1),
        ggplotly(p2))
```
## Highlight_key creates an pbject of class **crosstalk::SharedData**
 
```{r}
d <- highlight_key(exam_data)
p1<-ggplot(data=d, 
       aes(x=MATHS,
           y=ENGLISH))+
  geom_point(dotsize=1)+
  xlim(0, 100)+ylim(0, 100)+
  geom_abline(coef = c(0,1))

p2<-ggplot(data=d, 
       aes(x=MATHS,
           y=SCIENCE))+
  geom_point(dotsize=1)+
  xlim(0, 100)+ylim(0, 100)+
  geom_abline(coef = c(0,1))

subplot(ggplotly(p1),
        ggplotly(p2))
```

## **[DT](https://rstudio.github.io/DT/)**

```{r}
DT::datatable(exam_data)


```

## **Linked Brushing**: crosstalk method

-interactive selection between plot and table

```{r}
d <- highlight_key(exam_data)
p1<-ggplot(data=d, 
       aes(x=MATHS,
           y=ENGLISH))+
  geom_point(dotsize=1)+
  xlim(0, 100)+ylim(0, 100)+
  geom_abline(coef = c(0,1))

gg<-highlight(ggplotly(p1),"plotly_selected") 

crosstalk::bscols(gg,
                  DT::datatable(d),
                  widths = 5)

```


