---
title: " VAST Challenge 2021 -Mini-Challenge 3"
description: |
  A short description of the post.
author:
  - name: Linya Huang
    url: https://www.linkedin.com/in/linya-huang/
date: 07-03-2021
output:
  distill::distill_article:
    toc: true
    toc_depth: 6
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.retina = 3,
  echo = TRUE,
                      eval = TRUE,
                      message = FALSE,
                      warning=FALSE)
```

# 1.Introduction



# 2.Literature Review



# 3.Data Preprocessing and Exploratory

## 3.1 Import packages and social media stream data

```{r}
packages= c('raster','sf','clock','tmap',
            'tidyverse','data.table','lubridate',
            'textclean','tm','wordcloud','wordcloud2','text2vec',
            'topicmodels','tidytext','textmineR','quanteda',
            'BTM','textplot','concaveman')

for(p in packages){
  if(!require(p,character.only= T)){
    install.packages(p)
    }
  library(p, character.only = T)
}
```
```{r}
#install.packages("devtools")
library(devtools)
#install_github("cbail/textnets")
library(textnets)
```



```{r}
#read csv file
data_17_1830=read_csv("data/MC3/csv-1700-1830.csv")
data_1830_20=read_csv("data/MC3/csv-1831-2000.csv")
data_20_2130=read_csv("data/MC3/csv-2001-2131.csv")

#append data

data=rbindlist(list(data_17_1830,data_1830_20,data_20_2130))

#print head of data
head(data)

```
 

## 3.2 Change date type

```{r}

#timestamp in lubridate
data$timestamp <- ymd_hms(data$`date(yyyyMMddHHmmss)`)

```

## Data Exploratory
  
  •1. Number of post throughout the time period
  
  
```{r}
data$time_1min = cut(data$timestamp, breaks="1 min")
 
count <- data %>%
  group_by(type,time_1min) %>%
  summarise(count_of_posts= n_distinct(message))

count$time_1min=ymd_hms(count$time_1min)
#count$time_1min=format(count$time_1min,format = "%H:%M:%S") 


ggplot(count,aes(x=time_1min,y=count_of_posts,fill=type))+
         geom_bar(stat="identity",position="dodge")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  ggtitle("Total Number of Posts through the period")


```
```{r}
head(count[order(-count$count_of_posts),],10)
```
 
 With time interval at 1 minute, the number of posts trend is plotted as above. The peaks is observed during time 19:40-19:50. And the posts are also active around time 18:45:00 and 20:10:00.
 
 
 
## Text Data Preprocessing

```{r}


data$cleaned<-
  tolower(data$message)%>%   # transform all message to lower cases
  replace_contraction()%>%   #replace contractions with long form
  replace_word_elongation()%>% #remove the same letter (case insensitive) appears 3 times consecutively
  str_replace_all("[0-9]", "") %>% #removing numbers
  str_replace_all("(^[@#])|[[:punct:]]","")%>% #remove punctuations
  # leave # and @
  removeWords(stopwords("english"))%>% 
  str_squish()%>% #trim whitespace from a string 
  str_trim #removes whitespace from start and end of string


head(subset(data,select=c("message","cleaned")))


```



 
```{r}

#convert dataframe to corpus
docs <- Corpus(VectorSource(as.character(data$cleaned)))

inspect(docs[1:2])

# Create a document-term-matrix

dtm <- TermDocumentMatrix(docs) 
matrix <- as.matrix(dtm) 
words <- sort(rowSums(matrix),decreasing=TRUE) 

# words and frequency dataframe
df <- data.frame(word = names(words),freq=words)

#word cloud
wordcloud(words = df$word, freq = df$freq, min.freq = 5,           max.words=200, random.order=FALSE, rot.per=0.35,            colors=brewer.pal(8, "Dark2"))
wordcloud2(data=df, size=1.6, color='random-dark')

```


# 4.Data Visualization for Challenge Questions



#4.1 Using visual analytics, characterize the different types of content in the dataset. What distinguishes meaningful event reports from typical chatter from junk or spam? Please limit your answer to 8 images and 500 words.


To distinguish meaningful events related to the kidnap events happened in , firstly, I will perform topic modeling to identify the topic in the microblogs and apply LDAvis visualization for analysis.

```{r}
wordcorpus <- Corpus(VectorSource(as.character(data$cleaned)))  
dtm <- DocumentTermMatrix(wordcorpus,
                          control = list(
                            wordLengths=c(2, Inf),               # limit word length
                            bounds = list(global = c(5,Inf)),    # minimum word frequency
                            removeNumbers = TRUE,                #remove Numbers
                            weighting = weightTf,                #weighted term frequency
                            encoding = "UTF-8"))

rowTotals <- apply(dtm , 1, sum) #Find the sum of words in each Document
dtm.new   <- dtm[rowTotals> 0, ] #remove 0 dtm rows of matrix

topic=LDA(dtm.new,k=10,method="Gibbs",conrol=list(seed=2021,alpha=0.5,iter=200))
```
```{r}
terms(topic,5)
```

```{r}
#Tokenize data
tidytxtdata<- tidy(dtm)
                    
tidytxtdata <- tidytxtdata%>% #Remove the count column
  select(-count)
tidytxtdata <- tidytxtdata%>% #Change the column name 'term' to 'word' so that we can get rid of stopwords later
  rename(word = term)


#Remove stopwords
tidytxtdata <- tidytxtdata%>%
  anti_join(stop_words)


#Use the btm model
set.seed(321)
model <- BTM(tidytxtdata, k = 20, beta = 0.01, background = TRUE, iter = 500, trace = 100) #Run the model

topicterms <- terms(model, top_n = 10) #View the topics
topicterms
```
```{r}
library(textplot)
library(ggraph)
library(concaveman)
plot(model)
```



#4.2 Use visual analytics to represent and evaluate how the level of the risk to the public evolves over the course of the evening. Consider the potential consequences of the situation and the number of people who could be affected. Please limit your answer to 10 images and 1000 words.

```{r}
data_2<-data %>%
     group_by(author) %>%
     dplyr::summarise(all_posts=paste(cleaned, collapse=" "))

sotu_firsts_nouns <- PrepText(data_2, groupvar = "author", textvar = "all_posts", node_type = "groups", tokenizer = "words", pos = "nouns", remove_stop_words = FALSE, compound_nouns = TRUE)

# <- PrepTextSent(data_2, groupvar = "author", textvar = "cleaned", node_type = "groups", 
                                      #tokenizer = "words", sentiment_lexicon = "afinn", language = "english", udmodel_lang = #udmodel_lang, remove_numbers = NULL, compound_nouns = TRUE)

sotu_firsts_network <- CreateTextnet(sotu_firsts_nouns)

```
```{r}
#VisTextNet(sotu_firsts_network,label_degree_cut = 0)
VisTextNetD3(sotu_firsts_network)
```

```{r}
sotu_firsts_communities <- TextCommunities(sotu_firsts_network)
top_words_modularity_classes <- InterpretText(sotu_firsts_network, sotu_firsts_nouns)
text_centrality <- TextCentrality(sotu_firsts_network)
```


#4.3 If you were able to send a team of first responders to any single place, where would it be? Provide your rationale. How might your response be different if you had to respond to the events in real time rather than retrospectively? Please limit your answer to 8 images and 500 words


