---
title: "GeoVisual with R"
description: |
 In class exercise - Handling Geospatial Data with R and Visualizing Geospatial Data with R.
author:
  - name: Linya Huang
    url: https://www.linkedin.com/in/linya-huang/
date: 07-03-2021
output:
  distill::distill_article:
    toc: true
    toc_depth: 6
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message = FALSE)
```

# Install and Launch R packages

Insert an R code chunk either using the RStudio toolbar (the Insert button) or the keyboard shortcut Ctrl + Alt + I (Cmd + Option + I on macOS).


```{r, include=TRUE,message = FALSE}
packages= c('raster','sf','clock','tmap','tidyverse')

for(p in packages){
  if(!require(p,character.only= T)){
    install.packages(p)
    }
  library(p, character.only = T)
}
```

# Import **Raster** File


multiple bands - 3 different sublayers (RGB)

dimension - rows*columns

resolution - pixel size

crs - distance is in decimal, measurement will require scale manipulation

```{r}
ap<- raster("data/geospatial/MC2-tourist.tif")
ap
```
# Apply mapping packages in R - **tmap**

Plot Raster Layer

```{r}
bgmap<-ap

tmap_mode("plot") #interactive view

tm_shape(bgmap) +
  tm_raster(bgmap,
            legend.show = FALSE) # read as single band, not very useful in our case
```

```{r}
tmap_mode("plot") #interactive view

tm_shape(bgmap) +
  tm_rgb(bgmap, r=1, g=2, b=3,
         alpha=NA,
         saturation = 1,
         interpolate = TRUE,
         max.value = 255)
```
# Import Vector GIS Data File **st_read**

Geodetic CRS:  [WGS 84](https://en.wikipedia.org/wiki/World_Geodetic_System) -pay attention to this projection

```{r}
Ablia_st <- st_read(dsn="data/Geospatial",layer="Abila") # no need to add extension
```
# Import Aspatial Data 

timesamp filed is not in data-time format
id filed should be in factor data type

```{r}
gps <- read_csv("data/aspatial/gps.csv")
glimpse(gps) #nice tidyverse function
```

## convert date-time filed [clock](https://clock.r-lib.org/)

```{r}

gps$Timestamp <- date_time_parse(gps$Timestamp,
                                 zone="",
                                 format = "%m/%d/%Y %H:%M:%S")
gps$id <- as_factor(gps$id)
  
gpsl <- gps %>%
  filter(Timestamp < "2014-01-06 23:59:00")
  
```

```{r}

gpsl <- gps %>%
  filter(Timestamp < "2014-01-06 23:59:00")
  
```

```{r}

gps
  
```

# Convert aspatial data into a simple feature data frame

```{r}
gps_sf <- st_as_sf(gps,
                   coords = c("long","lat"),
                  crs= 4326)

gps_sf # geometry feature 
```

# Create Movement Path from GPS Points

use driver ID as unique identifiers

```{r}

gps_path <- gps_sf %>%
  group_by(id) %>%
  summarize(m=mean(Timestamp),  # for group by purpose , no meaning at all,just to summary
            do_union=FALSE) %>%
  st_cast("LINESTRING")

gps_path 
```

# Plot the gps Paths

```{r}
gps_path_selected <- gps_path %>%
  filter(id==1)

tmap_mode("view") # static mode
tm_shape(bgmap) +
  tm_rgb(bgmap, r=1, g=2, b=3,
         alpha=NA,
         saturation = 1,
         interpolate = TRUE,
         max.value = 255)+
  tm_shape(gps_path_selected)+
  tm_lines()
```

