---
title: "Network Visualization with R"
description: |
  Network Visualization with R
author:
  - name: Linya Huang
    url: https://www.linkedin.com/in/linya-huang/
date: 07-10-2021
output:
  distill::distill_article:
    toc: true
    toc_depth: 6
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.retina = 3,
  echo = TRUE,
                      eval = TRUE,
                      message = FALSE,
                      warning=FALSE)
```



# Import Data

```{r}
packages= c('igraph','tidygraph','ggraph','visNetwork',
            'lubridate','clock','tidyverse')

for(p in packages){
  if(!require(p,character.only= T)){
    install.packages(p)
    }
  library(p, character.only = T)
}
```

```{r}
GAStech_nodes <- read_csv("data/GAStech_email_node.csv")
GAStech_edges <- read_csv("data/GAStech_email_edge-v2.csv")

glimpse(GAStech_edges)
```
# Wrangling time


```{r}
GAStech_edges$SentDate <- dmy(GAStech_edges$SentDate)

GAStech_edges$Weekday = wday(GAStech_edges$SentDate,
                             label=TRUE,
                             abbr=FALSE)

glimpse(GAStech_edges)

```
# Wrangling attributes

Aggregate the individual by date, senders, receivers, main subject and day of the week.

```{r}
GAStech_edges_aggregated <- GAStech_edges %>%
  filter(MainSubject=="Work related") %>%
  group_by(source, target, Weekday) %>%
  summarise(Weight=n()) %>% #count 
  filter(source!=target)%>%  #remove archive email
  filter(Weight>1)%>%
  ungroup() 

glimpse(GAStech_edges_aggregated)
```
# Create tbl_graph() to build tidygraph date model

```{r}
GAStech_graph <- tbl_graph(nodes = GAStech_nodes, # id much match target edges
                           edges= GAStech_edges_aggregated,
                           directed=TRUE)

GAStech_graph 
```
```{r}
GAStech_graph <- tbl_graph(nodes = GAStech_nodes, # id much match target edges
                           edges= GAStech_edges_aggregated,
                           directed=TRUE)

GAStech_graph 
```

# Plot a basic network graph

__ggraph__
```{r}
ggraph(GAStech_graph)+
  geom_edge_link() +
  geom_node_point()
```

## remove background

```{r}
g<-ggraph(GAStech_graph)+
  geom_edge_link(aes()) +
  geom_node_point(aes())

g + theme_graph()
```
## change the color of plot


```{r}
g<-ggraph(GAStech_graph)+
  geom_edge_link(aes(color='grey50')) +
  geom_node_point(aes(color='grey40'))

g + theme_graph(background = "black")
```

## Fruchterman and Reingold layout

```{r}
g<-ggraph(GAStech_graph,layout='fr')+
  geom_edge_link(aes()) +
  geom_node_point(aes())

g + theme_graph()
```

## modify network nodes

```{r}
g<-ggraph(GAStech_graph,layout='nicely')+
  geom_edge_link(aes()) +
  geom_node_point(aes(colour=Department,
                      size=3))

g + theme_graph()
```

## modify edges

```{r}
g<-ggraph(GAStech_graph,layout='nicely')+
  geom_edge_link(aes(width=Weight),
                 alpha=0.2) +
  scale_edge_width(range=c(0.1,5))+
  geom_node_point(aes(colour=Department,
                      size=3))

g + theme_graph()
```


## creact facet graphs

how interaction change over weekdays
```{r}

set_graph_style()

g<-ggraph(GAStech_graph,layout='nicely')+
  geom_edge_link(aes(width=Weight),
                 alpha=0.2) +
  scale_edge_width(range=c(0.1,5))+
  geom_node_point(aes(colour=Department,
                      size=3))+
  theme(legend.position = 'bottom')

g + facet_edges(~Weekday)
```

high intensity on Tuesday bwteern Administration,Executive and Security department



```{r}

set_graph_style()

g<-ggraph(GAStech_graph,layout='nicely')+
  geom_edge_link(aes(width=Weight),
                 alpha=0.2) +
  scale_edge_width(range=c(0.1,5))+
  geom_node_point(aes(colour=Department,
                      size=3))
  

g + facet_nodes(~Department)+
  th_foreground(foreground = "grey80",
                border = TRUE)+
  theme(legend.position = 'bottom')
```


# Network Metrics Analysis

```{r}

g<-GAStech_graph %>%
  #mutate(betweeness_centrality=centrality_betweenness()) %>%
  ggraph(layout='fr')+
  geom_edge_link(aes(width=Weight),
                 alpha=0.2)+
  scale_edge_width(range = c(0.1,5))+
  geom_node_point(aes(colour=Department,
                      size=centrality_betweenness()))
  

g + theme_graph()


```

## Visualize community

```{r}

g<-GAStech_graph %>%
  mutate(community= as.factor(group_edge_betweenness(weight=Weight,directed = TRUE))) %>%
  ggraph(layout='fr')+
  geom_edge_link(aes(width=Weight),
                 alpha=0.2)+
  scale_edge_width(range = c(0.1,5))+
  geom_node_point(aes(colour=community))
  

g + theme_graph()

```

# Interactive Network Visualization

##Data Prep

```{r}
GAStech_edges_aggregated <- GAStech_edges %>%
  left_join(GAStech_nodes,by=c("sourceLabel"="label")) %>%
  rename(from=id) %>%
  left_join(GAStech_nodes,by=c("targetLabel"="label"))%>%
  rename(to=id) %>%
  filter(MainSubject=="Work related") %>%
  group_by(from,to)%>%
  summarise(weight=n())%>%
  filter(from!=to)%>%
  filter(weight>1)%>%
  ungroup()

GAStech_edges_aggregated
```
## __visNetwork__
```{r}
visNetwork(GAStech_nodes,
           GAStech_edges_aggregated) %>%
  visIgraphLayout(layout = "layout_with_fr")
```


## visual attributed- nodes
```{r}
GAStech_nodes <- GAStech_nodes %>%
  rename(group=Department)
```


```{r}
visNetwork(GAStech_nodes,
           GAStech_edges_aggregated) %>%
  visIgraphLayout(layout = "layout_with_fr") %>%
  visOptions(highlightNearest = TRUE,
             nodesIdSelection = TRUE)%>%
  visLegend() %>%
visLayout(randomSeed = 123) 
```



